Make any stupid plot you want

### 2021-02-09 draft/sections/intro.tex
rearranging, words

\subsection{What even is a data visualization library?}
Data visualizations originated as maps of places and stars \cite{friendlyBriefHistoryData2008}; direct graphical representations of physical items in space. While often decorative, to be useful these visualizations at least somewhat had to preserve some notion of relative distance between the objects being graphed. Byrne et al. deconstruct  visualizations as having either graphic or figurative components; graphic representations are maps from data values to visual encodings, while figurative representations have meaning due to their similarity in shape to external concepts \cite{byrneAcquiredCodesMeaning2016}. For example, the Du Bois data portraits in figure~\ref{fig:intro_dubois} are mostly graphic representations where quantitative variables are mapped to glyph segment length or area. In figure~\ref{fig:intro_dbc}, Du Bois combines a graphical representation where glyph size varies by population with a figurative representation of those glyphs as the countries the data is from.

@@

@@

set up: dubois

theorists:
bertin, munzner, mackinlay

talk about:
matplotlib arch paper, excel/matlab arch, 
vtk \& ggplot (compare/contrast, we're blending these things)

\subsection{Terminology}
\textcolor{blue}{Is currently C\&P'ed from my second exam so needs some messaging, but }
\begin{figure}[h!]
%%remake properly at some point
\includegraphics[width=\textwidth]{figures/intro/flowchart.png} 
\caption{To go from a dataset to a visualization, the data is subset based on a set of constraints (the invariant). The resulting subset becomes the components that are visualized, but the choice of visualization is dependent on the type and structure of the component variables.}
\label{fig:flowchart}
\end{figure}

\subsubsection{Data Type and Structure}
%% mention tidy/spivak first, set up that we use variable, record, key/index notation
%% define continuity in the math, connectivity in the data
%% topological spaces are frame in terms of continuity
% the implementation of $K$ embeds the connectivity
\label{sec:intro_data_structure}
\begin{figure}[h!]
 \includegraphics[width=\textwidth]{figures/intro/munzner_datatypes}
\caption{Keys are unique lookup values used to find individual observations in the dataset. Keys are positional references, and can be coordinates on a map or unique values such as a primary key in a database or a (time, latitude, longitude) index in a data cube. Image modified from a diagram from Munzner's \textit{Visualization Analysis and Design} \cite{munznerVisualizationAnalysisDesign2014}}
\label{fig:munzner_datatypes}
\end{figure}

@@

>>>>

\subsection{What even is a data visualization library?}
 Byrne et al. describe visualizations as having graphic and figurative components; graphic representations are maps from data values to visual encodings, while figurative representations have meaning due to their similarity in shape to external concepts \cite{byrneAcquiredCodesMeaning2016}. For example, the Du Bois data portraits in figure~\ref{fig:intro_dubois} are mostly graphic representations where quantitative variables are mapped to glyph segment length or area. Du Bois combines a graphical representation where glyph size varies by population with a figurative representation of those glyphs as the countries the data is from in figure~\ref{fig:intro_dbc}.

 Data visualizations originated as maps of places and stars \cite{friendlyBriefHistoryData2008}; direct graphical representations of physical items in space. While often decorative, to be useful these visualizations at least somewhat had to preserve some notion of relative distance between the objects being graphed, which Tufte termed the Lie principal\cite{tufteVisualDisplayQuantitative2001}.

@@

\begin{figure}[h!]
%%remake properly at some point
\includegraphics[width=\textwidth]{figures/intro/flowchart.png} 
\caption{To go from a dataset to a visualization, the data is subset based on a set of constraints (the invariant). The resulting subset becomes the components that are visualized, but the choice of visualization is dependent on the type and structure of the component variables.}
\label{fig:flowchart}
\end{figure}

@@

set up: dubois
\subsubsection{Data Type and Structure}
%% mention tidy/spivak first, set up that we use variable, record, key/index notation
%% define continuity in the math, connectivity in the data
%% topological spaces are frame in terms of continuity
% the implementation of $K$ embeds the connectivity
\label{sec:intro_data_structure}
\begin{figure}[h!]
 \includegraphics[width=\textwidth]{figures/intro/munzner_datatypes}
\caption{Keys are unique lookup values used to find individual observations in the dataset. Keys are positional references, and can be coordinates on a map or unique values such as a primary key in a database or a (time, latitude, longitude) index in a data cube. Image modified from a diagram from Munzner's \textit{Visualization Analysis and Design} \cite{munznerVisualizationAnalysisDesign2014}}
\label{fig:munzner_datatypes}
\end{figure}

@@

\subsection{How do other visualization libraries tackle this?}
matplotlib arch paper, excel/matlab arch, 
vtk \& ggplot (compare/contrast, we're blending these things)
